How is experience transformed into memory?

### 2018-08-28 paper/main.tex
few minor edits and updated pdf

Previous work suggests that the brain may represent our experiences as ``events'' with different time scales in a nested hierarchy ~\citep{LernEtal11, ChenEtal17, BaldEtal17}. In this next analysis we asked: what brain regions mirror the event-like correlational structure of our topic-based video model during viewing? We used a searchlight procedure (see Methods for details) to identify regions whose timepoint-by-timepoint correlational structure matched the correlational structure of our model. This analysis revealed a network of regions including bilateral frontal and cingulate cortex (Fig. ~\ref{fig:brainz}), suggesting that these regions may play a role in maintaining information relevant to the narrative trajectory of the video.

While each participant remembered roughly the same sequence of events, there were subtle idiosyncracies in the way each individual described the events. For example, the event-by-event similarity structure (i.e. the off-diagonal values of the recall correlation matrices) differed across individuals (Supp. Fig.~\ref{fig:corrmats}). We hypothesized that idiosyncracies in the recall dynamics across subjects might be related to the temporal structure of brain networks representing the events during video viewing. In this analysis, we identified brain areas whose responses at the time of viewing exhibited the same (segmented) idiosyncratic dynamics that later appeared in that participants’ recalls. Using a procedure similar to the approach above (see Methods for details), we searched for regions whose correlational structure during viewing was related to the idiosyncratic structure of the participant's recall model. The resulting map suggests that the correlational structure of a network of brain regions including the ventromedial prefrontal cortex, anterior cingulate, and right medial temporal lobe mirror the idiosyncratic structure of participants' recalls (Fig. ~\ref{fig:brainz}).

@@

\subsubsection{Vector field analysis}
To quantify the flow of recall from event to event, we performed a vector field analysis.  We tiled the 2D topic space (x, y: -6 to 6 by .25) with an evenly spaced grid. For each grid point, we drew a circle around the point (radius=.5). Then, we tested whether any line segments (formed by event recall transitions) passed through this area of the topic space.  For example, say that a participant transitioned from recalling event 2 to event 3. These 2 recall events correspond to 2 points in topic space, and connecting them forms a line segment. We collected all line segments that passed through a given section of topic space (collapsing across participants). To plot the average direction of the line segments (i.e., the arrows for each grid point in ~\ref{fig:trajectory}B), we converted each of them to unit vectors and then averaged. For grid points where the direction was consistent (across all participants contributing to that point), the length of the arrow approaches 1, whereas if the direction was random the length of the arrow approaches 0. Lastly, we converted each unit vector to an angle (in radians) by taking the inverse tangent of the x, y components of the vector. To test whether the distribution of angles was significantly non-uniform (i.e., displayed a preferred direction across participants), we performed a Rayleigh test on the angles ($p<.001$, FDR-corrected at $p<.05$ using Benjamani-Hochberg procedure). Arrows where the Rayleigh test was significant are displayed in color (the darker the blue the more significant) while non-significant tests are displayed in gray with lower opacity.

@@

\pagebreak
\begin{center}
\textbf{\large Supplemental Materials: Capturing the shape of naturalistic experiences and how we remember them}
\end{center}
\setcounter{equation}{0}
\setcounter{figure}{0}
\setcounter{table}{0}
\setcounter{page}{1}
\setcounter{section}{0}
\makeatletter
\renewcommand{\theequation}{S\arabic{equation}}
\renewcommand{\thefigure}{S\arabic{figure}}
\renewcommand{\bibnumfmt}[1]{[S#1]}
\renewcommand{\citenumfont}[1]{S#1}

>>>>

Previous work suggests that the brain may represent our experiences as ``events'' with different time scales in a nested hierarchy ~\citep{LernEtal11, ChenEtal17, BaldEtal17}. In this next analysis we asked: what brain regions mirror the event-like correlational structure of our topic-based video model during viewing? We used a searchlight procedure (see Methods for details) to identify regions whose timepoint-by-timepoint correlational structure matched the correlational structure of our model. This analysis revealed a network of regions including bilateral frontal and cingulate cortex (Fig.~\ref{fig:brainz}), suggesting that these regions may play a role in maintaining information relevant to the narrative trajectory of the video.

While each participant remembered roughly the same sequence of events, there were subtle idiosyncracies in the way each individual described the events. For example, the event-by-event similarity structure (i.e. the off-diagonal values of the recall correlation matrices) differed across individuals (Supp. Fig.~\ref{fig:corrmats}). We hypothesized that idiosyncracies in the recall dynamics across subjects might be related to the temporal structure of brain networks representing the events during video viewing. In this analysis, we identified brain areas whose responses at the time of viewing exhibited the same (segmented) idiosyncratic dynamics that later appeared in that participants’ recalls. Using a procedure similar to the approach above (see Methods for details), we searched for regions whose correlational structure during viewing was related to the idiosyncratic structure of the participant's recall model. The resulting map suggests that the correlational structure of a network of brain regions including the ventromedial prefrontal cortex, anterior cingulate, and right medial temporal lobe mirror the idiosyncratic structure of participants' recalls (Fig.~\ref{fig:brainz}).

@@

\subsubsection{Vector field analysis}
To quantify the flow of recall from event to event, we performed a vector field analysis.  We tiled the 2D topic space (x, y: -6 to 6 by .25) with an evenly spaced grid. For each grid point, we drew a circle around the point (radius=.5). Then, we tested whether any line segments (formed by event recall transitions) passed through this area of the topic space.  For example, say that a participant transitioned from recalling event 2 to event 3. These 2 recall events correspond to 2 points in topic space, and connecting them forms a line segment. We collected all line segments that passed through a given section of topic space (collapsing across participants). To plot the average direction of the line segments (i.e., the arrows for each grid point in Fig.~\ref{fig:trajectory}B), we converted each of them to unit vectors and then averaged. For grid points where the direction was consistent (across all participants contributing to that point), the length of the arrow approaches 1, whereas if the direction was random the length of the arrow approaches 0. Lastly, we converted each unit vector to an angle (in radians) by taking the inverse tangent of the x, y components of the vector. To test whether the distribution of angles was significantly non-uniform (i.e., displayed a preferred direction across participants), we performed a Rayleigh test on the angles ($p<.001$, FDR-corrected at $p<.05$ using Benjamani-Hochberg procedure). Arrows where the Rayleigh test was significant are displayed in color (the darker the blue the more significant) while non-significant tests are displayed in gray with lower opacity.

@@

\pagebreak
\begin{center}
\textbf{\large Supplemental Materials: How is experience transformed into memory?}
\end{center}
\setcounter{equation}{0}
\setcounter{figure}{0}
\setcounter{table}{0}
\setcounter{page}{1}
\setcounter{section}{0}
\makeatletter
\renewcommand{\theequation}{S\arabic{equation}}
\renewcommand{\thefigure}{S\arabic{figure}}
\renewcommand{\bibnumfmt}[1]{[S#1]}
\renewcommand{\citenumfont}[1]{S#1}
